---
title: "Seurat Clustering"
output: html_notebook
---

---
editor: 
  markdown: 
    wrap: 72
---

## Setup the Seurat Object

For this project, I will analyze the Peripheral Blood Mononuclear Cells from the 10X Genomics. There are 2,700 single cells that were sequenced on the Illumina NextSeq 500.

Now, I will import the libraries that I will use to perform the clustering.

```{r}
# import the libraries
library(dplyr)
library(Seurat)
library(patchwork)
```

Now that we have the libraries, we can read the dataset.

```{r}
# read the dataset
pbmc.data <- Read10X(data.dir = "pbmc3k_filtered_gene_bc_matrices/filtered_gene_bc_matrices/hg19/")
```

We want our dataset to be in Seurat object. But, the pbmc.data is in matrix format.

```{r}
# check the dataset
class(pbmc.data)
```

So, the dataset has been read and saved in the pbmc.data variable but the data type is matrix. Now, we will create Seurat object from the dataset.

```{r}
# Create the Seurat Object
pbmc <- CreateSeuratObject(counts = pbmc.data, project = 'pbmc3k', min.cells = 3, min.features = 200)
```

```{r}
# visualize the dataset
pbmc.data[c("CD3D", "TCL1A", "MS4A1"), 1:8]
```

```{r}
dense.size <- object.size(as.matrix(pbmc.data))
dense.size
```

```{r}
sparse.size <- object.size(pbmc.data)
sparse.size
```

```{r}
dense.size / sparse.size
```

# Standard pre-processing workflow

These represent the selection and filtration of cells based on QC metrics, data normalization and scaling, and the detection of highly variable features.

## QC and selecting cells for further analysis

In this step, we will exluce some of the unnecesseray data from the datast, so ensure proper quality control.

```{r}
pbmc[["percent.mt"]] <- PercentageFeatureSet(pbmc, pattern = "^MT-")
```

```{r}
VlnPlot(pbmc, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```
